Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2017-03-26 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Completed Future research Chapter

\chapter{Future Research Needs}

This study examined a very limited set of conditions as a means to provide a baseline for future research efforts that would expand on the results from this study in order more completely address the National Academy of Science ...   natural repeatability of fire patterns.   Since the National Academy of Science report was issued in 2009, the  the U.S. National Institute of Justice, Fire and Arson Investigation Technology Working Group has identified 5 areas where scientific research is needed: (1) understanding of the effects of ventilation on fire damage and patterns; (2) characterization of electrical system response as a means to study fire progression; (3) repeatability and reproducibility of test measurements of large-scale structure fires, (4) materials property data inputs for accurate computer models, and (5) evaluation of incident heat flux profiles to walls and neighboring items in support of fire model validation~\cite{NIJ:2016}.  The areas suggested for future research in this study involve 4 of the 5 topic areas from the NIJ working group.

Pre-flashover experiments in compartments using well characterized burner fires should be conducted with a broad range of heat release rates and varied burn sizes to 
to determine if the repeatability results from this study can be replicated.

The flow field between the burner and the adjacent wall or walls should have further research conducted to better understand the movement of the flame and the plume.  The  when the fire was adjacent to the wall or a corner should receive further examination.   
distance from the wall

degradation of gypsum board full scale

Use of black paint

Calcination depth

Larger variation of heat release rates

Realistic fuels

duration

lessons learned  cameras   black paint

corners

While the results from this project will provide some insight into the repeatability of pre-flashover fire patterns, a significant amount of research examining other fuels, wall coverings, the impacts of compartmentation, ventilation, and post-flashover conditions, are just a few of the parameters that need to be examined to develop a more complete understanding of fire patterns.

Uncertainty and Modeling

@@

Fire investigators, however, examine fire scenes where the first item ignited may have involved a piece of furniture, an appliance, or even a car.   These items are three dimensional and the fuel could be emitted from multiple surfaces at any given time in the fire event.  Further thermoplastics are commonly used in the construction of these items, with furniture being a prevalent item.  As the thermoplastic is heated it melts, flows, drips, and burns.   As a result the flame geometry generated by a piece of furniture may not be representative of the flame geometry characterized by the studies using gas burners and pool fires.  In this study even using a block of polyurethane foam with a simple geometry led to increased uncertainty.

>>>>

\chapter{Future Research}

This study examined a very limited set of conditions as a means to provide a baseline for future research efforts that would expand on the results from this study in order more completely address the National Academy of Science ...   natural repeatability of fire patterns~\cite{Forensic:2009}.   Since the National Academy of Science report was issued in 2009, the  the U.S. National Institute of Justice, Fire and Arson Investigation Technology Working Group has identified 5 areas where scientific research is needed: (1) understanding of the effects of ventilation on fire damage and patterns; (2) characterization of electrical system response as a means to study fire progression; (3) repeatability and reproducibility of test measurements of large-scale structure fires, (4) materials property data inputs for accurate computer models, and (5) evaluation of incident heat flux profiles to walls and neighboring items in support of fire model validation~\cite{NIJ:2016}.  The areas suggested for future research in this study involve 4 of the 5 topic areas from the NIJ working group.

\section{Experiments}

While the results from this project provided a baseline regarding the repeatability of pre-flashover fire patterns, a significant amount of research examining other fuels, wall coverings, the impacts of compartmentation, ventilation, and post-flashover conditions, are just a few of the parameters that need to be examined to develop a more complete understanding of fire patterns.

Pre-flashover experiments in compartments using well characterized burner fires should be conducted with a broad range of heat release rates, burn durations, and varied burner sizes to determine if the repeatability results from this study can be replicated.

The flow field between the burner and the adjacent wall or walls should have further research conducted to better understand the movement of the flame and the plume.  This may also lead to improved information on the side to side motion of the flame when the burner was against the wall.

Research on fire patterns also needs to be conducted with fuels similar to those that the fire investigators will encounter in the field, such as furnishings and appliances. 
These fuels release fuel from 3-dimensional surfaces, as opposed to a 2-dimensional burner surface,  In addition, many furnishing items are composed of thermoplastic materials which may melt, flow and drip, potentially spreading the fire and changing the fire pattern during the course of the burn.

There is a great need to understand the impact of ventilation on fire patterns within a structure.  Fires that stop pre-flashover as well as fires that burn post-flashover need to be studied.  The structure itself will have an impact on the movement and flow of both fresh air to the fire as well as containing and channeling the movement of combustion products away from the fire, both of which can affect the development of the fire pattern.  Post-flashover fires are particularly challenging for fire investigators as the potential is high that significant fire damage may be generated remote from the area of origin.

There is a need for continued research on materials used in construction and other furnishings (first items ignited).  Since the start of this study, a few notable items have changed that may challenge the investigators.  The gypsum wall board used in this study is considered to be regular gypsum wall board, it was widely used and would be what was readily available at construction material supply stores across the U.S.  Today, the regular gypsum board has been replaced on store shelves by a less dense version of 12.7~mm thick gypsum wall board.  A finished lightweight gypsum wall board panel has 30~\% less mass than the regular gypsum wall board that was the de facto construction standard just a few years ago.  How the lightweight gypsum wall board degrades when exposed to fire has not been quantified.

\section{Modeling}

@@

Fire investigators, however, examine fire scenes where the first item ignited may have involved a piece of furniture, an appliance, or even a car.   These items are three dimensional and the fuel could be emitted from multiple surfaces at any given time in the fire event.  Further thermoplastics are commonly used in the construction of these items, with furniture being a prevalent item.  As the thermoplastic is heated it melts, flows, drips, and burns.   As a result the flame geometry generated by a piece of furniture may not be representative of the flame geometry characterized by the studies using gas burners and pool fires.  In this study even using a block of polyurethane foam with a simple geometry led to increased uncertainty of the peak heat release rate and decreased fire pattern repeatability.

While all of the experimental studies recommended in the previous section may support model validation studies, it is important that a specific validation study be conducted with fuels normally found in real fire investigations to determine the applicability of various numerical methods under realistic fire conditions.
